Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-11 GWLDCI.tex
proof read through statement of 3.17

The set of propagators $P= \{p,q,r,s \}$ defines an exact subdiagram of $W$, with $V(P) = \{1,2,3,4,5,8,9\}$.  The corresponding subgraph $t$ of $\tau(W)$ has vertex set $\{1, 2,3, 4, 8\}$. Note that we don't include vertex 5 because it corresponds to the edge in $W$ that connects vertices 5 and 6, but 6 does not belong to $V(P)$. Vertex 9 is excluded for similar reasons.

@@

The key point to note is that the subgraph $t$ depends on how the propagators of $P$ sit inside $[n]$, not on how they sit in $W_P$ itself.  In particular, $t$ is not the subgraph of $\tau(W)$ consisting only of edges corresponding to propagators of $P$ (as it may also include some of the outer edges of $\tau(W)$), nor is it the polygon dissection of $(P,V(P))$ viewed as a Wilson loop diagram in its own right (which would have vertex set equal to $V(P)$).
\end{eg}

@@

Each of these $m-3+j$ internal edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ internal edges.

The graph $t$ has $j$ edges that come from propagators, and $m - 2c$ edges that come from the boundary polygon of $\tau(W)$. Since $t$ has $m-c$ vertices, it has $m-c$ external edges, of which $c$ come from propagators. Therefore, of the $j$ edges of $t$ that come from propagators, $j-c$ are internal to the connected component. Therefore, $t$ is a triangulated piece.

@@

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally. Recall from Definition~\ref{VPropdfn} that the propagator flat $F(P)$ of $P \subseteq \cP$ denotes the set $V(P^c)^c$ of vertices that do not support any propagators not in $P$.

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $F(P^c)$ is a dependent flat. \note{Should we emphasize that all these $P^c$ are wrt $W$ not the subdiagram?}
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate}
\end{lem}

>>>>

The set of propagators $P= \{p,q,r,s \}$ defines an exact subdiagram of $W$, with $V(P) = \{1,2,3,4,5,8,9\}$.  The corresponding subgraph $t$ of $\tau(W)$ has vertex set $\{1, 2,3, 4, 8\}$. Note that we include vertex 3 of $\tau(W)$ because the two endpoints of edge 3 in $W$ (namely the vertices 3 and 4) are both in $V(P)$. However, we don't include vertex 5 of $\tau(W)$ because only of its endpoints in $W$ (5 and 6) are in $V(P)$, namely 5. The same argument applies to vertices 7, 9 and 10. Vertex 6 is excluded because none of its endpoints lie in $V(P)$.

@@

The key point to note is that the subgraph $t$ depends on how the propagators of $P$ sit inside $[n]$, not on how they sit in $W_P$, which would have an outer edge between vertices 8 and 1, one between 2 and 4, and another between 4 and 8, but would not include vertex 3.  In particular, $t$ is not the subgraph of $\tau(W)$ consisting only of edges corresponding to propagators of $P$ (as it may also include some of the outer edges of $\tau(W)$), nor is it the polygon dissection of $(P,V(P))$ viewed as a Wilson loop diagram in its own right (which would have vertex set equal to $V(P)$).
\end{eg}

@@

Each of these $m-3+j$ \hlfix{internal}{diagonal? chord?} edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose $|P|>1$.  Let $j=|P|$, $m= j+3=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ \hlfix{internal}{here, internal means that it doesn't define the outer face of $t$ in the induced embedding. Different use of the word than above?} edges.

The graph $t$ has $j$ edges that come from propagators, and $m - 2c$ edges that come from the boundary polygon of $\tau(W)$. Since $t$ has $m-c$ vertices, it has $m-c$ \hlfix{external}{means defines outer face, not just and outer edge of $\tau(W)$} edges, of which $c$ come from propagators. Therefore, of the $j$ edges of $t$ that come from propagators, $j-c$ are internal to the connected component. Therefore, $t$ is a triangulated piece.

@@

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally. Recall from Definition~\ref{VPropdfn} that the propagator flat $F(P)$ of $P \subseteq \cP$ denotes the set $V(P^c)^c$ of vertices that \st{do not support any propagators not in $P$.}\hlfix{only support propagators in $P$ (or no propagators)}{too many negatives. Can we use this instead?}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram of $W$, then $F(P^c)$ is a dependent flat in $W$. \note{Should we emphasize that all these $P^c$ are wrt $W$ not the subdiagram?} \sanote{yes. done}
\item If $(P,V(P))$ is a maximal exact subdiagram of $W$, then $F(P^c)$ is a cyclic flat in $W$.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate} \sanote{end here}
\end{lem}
